### 2019-04-01 papers/composite_inference/draft.tex
Changes last week of March 2019

\documentclass[english]{scrartcl}

\usepackage{fullpage}
\usepackage{amstext,amssymb,amsmath,amsthm}
\usepackage{graphicx,subfigure}
\usepackage{multirow}
\usepackage{url}
\usepackage{hyperref}
\usepackage{color}

\graphicspath{{.}{pics/}}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{proposition}[theorem]{Proposition}

\title{Composite Bayesian inference}
%\date{}
\author{Alexis Roche\thanks{\url{alexis.roche@centraliens.net}}}

\def\x{{\mathbf{x}}}
\def\y{{\mathbf{y}}}
\newcommand{\blambda}{{\boldsymbol{\lambda}}}
\newcommand{\bell}{{\boldsymbol{\ell}}}
\newcommand{\E}{\mathbb{E}}
%\newcommand{\matphi}{\boldsymbol{\Phi}} 
%\def\p{{\bar{\mathbf{p}}}}
%\def\q{{\bar{\mathbf{q}}}}

\begin{document}

\maketitle

\begin{abstract}
We revisit and generalize the concept of composite likelihood as a method to make a probabilistic inference by aggregation of multiple Bayesian agents. The resulting class of predictive models, which we call ``composite Bayesian'', is a middle way between purely generative and purely discriminative models. This perspective gives insight to choose the weights associated with composite likelihood, either {\em a priori} or via learning: in the latter case, they may be tuned so as to maximize training likelihood, yielding an easy-to-solve convex problem. We argue that composite Bayesian inference trades off between interpretability and prediction performance, both of which are crucial to many artificial intelligence tasks.
\end{abstract}

\section{Introduction}
\label{sec:intro}

Textbook statistical inference (frequentist or Bayesian) rests upon the existence of a probabilistic data-generating model that is both empirically valid and computationally tractable. Because this double requirement may be very challenging for multidimensional data, other inference models have been developed in applied science: deliberately misspecified generative models, as in quasi-likelihood \cite{White-82,Walker-13} or na\"ive Bayes \cite{Ng-01} methods; data compression models as in minimum description length \cite{Grunwald-07}; and discriminative models\footnote{A {\em discriminative} model is a parametric family that describes the distribution of a target variable conditional on the data, in contrast with a {\em generative} model, which describes the conditional distribution of the data given the target variable.}, which currently dominate the field of artificial intelligence (AI) and typically require supervised learning on large datasets -- these include many classical (shallow) learning \cite{Ho-95,BergerA-96,Vapnik-00,Rasmussen-06} and deep learning \cite{Lecun-15,Goodfellow-16} techniques (with the exception of deep belief networks \cite{Hinton-06}).

Discriminative models, however, lack the notion of {\em likelihood}: they cannot assess whether the data is consistent with their own prediction, however accurate, and in this sense do not have the ability to ``justify'' themselves. Consider as a straightforward example the problem of deciding whether an object is a sauce pan or a frying pan based on depth. A discriminative model can learn a threshold that correctly classifies most pans in practice, but is unable to see that, say, a blender does not match either category. The impossibility of such confirmatory analysis strongly limits interpretation if outlier inputs are to be expected.

There is growing awareness that AI should be interpretable in life-impacting applications~\cite{Molnar-18}, where it is (and perhaps should remain) confined to a supportive role in human-driven decision making processes. For instance, in medical diagnosis, automated disease predictions should be corroborated by individualized findings to be communicable to clinicians. In other words, there is no interest for ``black boxes'' but for AI systems that can justify their opinions just like human experts. On the other hand, clinicians have long used reference range analysis for diagnosis, which stems from simple univariate generative models.

%Imagine a doctor telling a patient: ``Sorry, sir, you have Alzheimer's disease because this fancy AI software said so''. Unless the said software is 100\% reliable, which is practically impossible, this would obviously be unacceptable. Acceptable would be to corroborate the diagnosis with some key quantitative empirical observations: low memory test scores, hippocampal atrophy, etc. In other words, inference has to come with some insight. An AI system is but one expert relying on hypotheses which may be invalid, as any expert, and is therefore prone to error. A typical error source is to train a classifier on a subset of classes from the real world. We cannot avoid inference errors but we can safeguard against them by being able to interrogate the system: why do you think what you think? To achieve that, the system should be capable of {\em introspection.

%Vapnik: ``one should solve the classification problem directly and never solve a more general problem as an intermediate step''. True if the goal is classification only. Wrong if introspection is needed because the problem is then more general in nature.

%Human inference does not work in a pre-determined universe of ``causes''. We confront theories with reality to evaluate how likely they are, while knowing that none of the theories considered so far may be ``right''. In fact, we are not looking for the truth, but for a conving enough theory.

%Composite likelihood seen as a pool of experts comes with a potential introspection mechanism: first sort experts by decreasing influence on the predictive distribution, see how likely the different classes are to the most influential experts.

A related limitation of discriminative models is that they are not suitable for unsupervised learning or on-the-fly parameter estimation because they treat the data and the model parameters as {\em marginally independent} variables, meaning that the data conveys no information about the parameters unless the target variable is observed. This is illustrated in Figure~\ref{fig:graph_comparison} by the respective directed graphs representing generative and discriminative models. For the same basic reason, supervised learning in a discriminative model is statistically less efficient than in a generative model spanning the same family of posteriors, hence it requires more training data to achieve optimal performance \cite{Ng-01}.

%Overall, pure discriminative models are of little use outside the context of big labeled data.

\begin{figure}[!ht]
\begin{center}
\subfigure[Generative model]{\includegraphics[width=.25\textwidth]{generative.pdf}\label{fig:generative}}
\hspace*{.2\textwidth}
\subfigure[Discriminative model]{\includegraphics[width=.25\textwidth]{discriminative.pdf}\label{fig:discriminative}}
\caption{Directed graphs representing generative and discriminative models, where $X$, $Y$ and $\theta$ respectively denote the target variable, the data, and the model parameters. Note the marginal independence of data and parameters in the discriminative model.}
\label{fig:graph_comparison}
\end{center}
\end{figure}

This note advocates probabilistic opinion pooling \cite{Genest-86} as a natural way to merge discriminative and generative modeling approaches in order to make predictions that are both computationally efficient and interpretable, resulting in a class of hybrid models. The key idea is to combine multiple low-dimensional generative models corresponding to different pieces of information extracted from the input data. Each such model acts as an isolated ``agent'' that uses a single feature to express an opinion in the form of a likelihood function of the target variable. The agent opinions may then be aggregated into a unique predictive probability distribution analogous to a Bayesian posterior. This may be understood as a probabilistic version of boosting, or as a {\em divide and conquer} approximation to the intractable Bayesian posterior.

%Importantly, predictions made in such a way are interpretable owing to their underlying generative nature.

When choosing the aggregation method as a log-linear pool, the predictive distribution turns out to be proportional to a quantity known in computational statistics as composite likelihood (CL). As discussed in \cite{Varin-11} and the references therein, CL was developed as a surrogate of traditional likelihood for parameter estimation. While maximum CL does not inherit the general property of maximum likelihood to achieve asymptotical minimum variance, it is asymptotically consistent under mild conditions \cite{Xu-11} and may offer an excellent trade-off between computational and statistical efficiency.

With the opinion pooling interpretation in mind, the weights assigned to the different features may be optimized for prediction performance in a typical discriminative learning scenario. As we will see, this strategy amounts to training a maximum entropy classifier using the feature-based log-likelihoods as basis functions. The likelihood functions themselves may need to be pre-trained, therefore the composite inference approach typically involves a double training scheme: a generative training stage (to learn the feature-based likelihood parameters) followed by a discriminative training stage (to learn the feature weights in aggregation).

Technical details are given in the remainder. We warn the reader that slightly abusive mathematical notations are used deliberately for the sake of clarity in places where we think that they are unambiguous.

\section{Composite likelihood as opinion pooling}
\label{sec:log_pool}

Let $\mathbf{Y}$ an observable multivariate random variable with sampling distribution $p(\y|x)$ conditional on some unobserved variable of interest, $X\in{\cal X}$, where ${\cal X}$ is assumed to be a finite set for simplicity. Given an experimental outcome $\y$, the likelihood is the sampling distribution evaluated at $\y$, seen as a function of $x$:
$$
L(x) = p(\y|x)
.
$$

This requires a plausible generative model which, for complex data, may be out of reach or involve too many nuisance parameters. A natural workaround known as data reduction is to extract some lower-dimensional representation $z(\y)\sim f(z|x)$ using a many-to-one mapping, and consider the potentially more convenient likelihood function:
$$
\ell(x) = f(z|x)
.
$$

Substituting $L(x)$ with $\ell(x)$ boils down to restricting the feature space, thereby  ``delegating'' statistical inference to an ``agent'' provided with partial information. While it is valid for such an agent observing~$z$ only to consider $\ell(x)$ as the likelihood function of the problem, the drawback is that $\ell(x)$ might yield too vague a prediction of~$X$ due to the information loss incurred by data reduction. To make the trick statistically more efficient, we may extract several features, $z_i(\y)$ for $i=1,2,\ldots,n$, and try to combine the likelihood functions $\ell_i(x) = f(z_i|x)$ that they elicit.

If we see the likelihoods as Bayesian posterior distributions corresponding to {\em uniform} priors, this is a problem of probabilistic opinion aggregation from possibly redundant sources, for which several methods exist in the literature \cite{Tarantola-82,Genest-86,Garg-04,Allard-12}. One that is appealing as it tends to produce single peaked distributions is the (generalized) log-linear opinion pool:
\begin{equation}
\label{eq:log_pool}
p_\blambda(x|\y) \propto \pi(x) \prod_{i=1}^n f(z_i|x)^{\lambda_i},
\end{equation} 
where $\pi(x)$ is a prior distribution, and $\blambda=(\lambda_1,\ldots,\lambda_n)$ is a vector of weights. Note that it is possible to have the weights depend on the data~$\y$, but this would make the method considerably more complex and less interpretable. Negative weights should be ruled out to warrant a natural monotonicity property: if all agents agree that an outcome $x_1$ is less likely than an outcome $x_2$, this should be reflected by the consensus probabilities, {\em i.e.}, $p_\blambda(x_1|\y)\leq p_\blambda(x_2|\y)$.

%This condition holds whenever all weights are positive ($\lambda\succeq 0$), assuming that the agent opinions are upper bounded. Negative weights should be ruled out to warrant a natural monotonicity property: if all agents agree that an outcome $x_1$ is less likely than an outcome $x_2$, then the consensus probabilities should always satisfy $p_\lambda(x_1|\y)\leq p_\lambda(x_2|\y)$. Note that it is possible to have the weights depend on the data~$y$, but this would make the method considerably more complex and less interpretable.

%Strictly positive weights guarantee the so-called 0/1 forcing property, that is, if an hypothesis~$x$ has zero likelihood according to at least one agent, then its consensus probability vanishes too.

% Not clear yet as to what a negative weight could mean!

\section{Composite Bayes rule}
\label{sec:bayes_rule}

The log-linear pool~(\ref{eq:log_pool}) bears a striking similarity to Bayes rule as it yields the form: 
$$
p_\blambda(x|\y)\propto \pi(x) L^c_\blambda(x),
$$
where the distribution $\pi(x)$ plays the role of a prior, and the function:
\begin{equation}
\label{eq:comp_lik}
L^c_\blambda(x) \equiv \prod_{i=1}^n \ell_i (x)^{\lambda_i}
\end{equation} 
plays that of a likelihood function. The expression~(\ref{eq:comp_lik}) happens to be known in computational statistics as a {\em marginal composite likelihood} \cite{Varin-11}. The slightly more general {\em conditional composite likelihood} form can be derived in the same way as above by conditioning all probabilities on confounding variables, see Appendix~\ref{app:conditional}. The computational advantage of CL over a genuine likelihood function stems from the fact that it is more efficient to evaluate the marginal distributions of each feature than the joint distribution of all features.

CL shares a convenient factorized form with the likelihood derived under the assumption of mutual feature independence, often referred to as {\em na\"ive Bayes} in the machine learning literature, which corresponds to the special case of unitary weights, $\blambda\equiv 1$. While the CL derivation does not assume feature independence, it yields an alternative interpretation of na\"ive Bayes as a general feature aggregation method, however unitary CL weights do not guarantee optimal prediction performance.

\section{Composite likelihood calibration}
\label{sec:calibration}

CL involves two types of parameters:
\begin{enumerate}
    \item ``generative'' parameters, {\em i.e.}, the parameters of the feature generation models $f(z_i|x)$; 
    \item ``discriminative'' parameters, {\em i.e.},  the weights~$\blambda=(\lambda_1,\ldots,\lambda_n)$ assigned to the feature-based likelihood functions in~(\ref{eq:log_pool}).
\end{enumerate}

If training data is available, the generative parameters may be set in a first stage using standard estimation techniques. Otherwise, they may be considered as nuisance parameters and eliminated at prediction time by applying to composite likelihood one of the same techniques as for traditional likelihood \cite{Berger-99}. We focus in the sequel on tuning the composite weights.

\subsection{Agnostic weights}

In the absence of knowledge, uniform weights should be chosen under a rule of indifference. CL is then a scaled version of na\"ive Bayes likelihood, the unique weight value being irrelevant to the maximum CL estimator (MCLE). To get meaningful predictive probabilities, it may be tuned empirically so as to best adjust the pseudo posterior variance matrix to the asymptotic MCLE variance matrix \cite{Pauli-11}, or via a close-in-spirit curvature adjustment \cite{Ribatet-12}, as proposed in previous attempts at Bayesian inference from composite likelihood.

% in attempts to match frequentist and Bayesian notions of uncertainty.

Alternatively, a simple agnostic recommendation we believe to be useful for composite weights is to sum up to one. This is motivated by the opinion pooling perspective and a characterization theorem \cite{Genest-86,Genest-86b}: the only {\em externally Bayesian} pooling operator (for which the consensus probability of an outcome only depends on the agent probabilities of the same outcome) is the log-linear opinion pool with unit sum weights. The external Bayesian  property essentially means that the consensus should not change if an agent opinion is taken into account by the other agents as opposed to being stated independently. The log-linear pool with unit sum weights also happens to be an optimal {\em compromise} in the sense that it minimizes the average inclusive Kullback-Leibler (KL) divergence to the agent opinions \cite{Garg-04}.

%While other divergence measures lead to other optimal pooling operators (for instance, the exclusive KL divergence leads to the linear pool), they all enforce unit sum weights, as is natural in the search for a best compromise.

Unit sum weights, however, implicitly assume maximum redundancy between features. Consequently, they tend to produce conservative predictive probabilities (see Appendix~\ref{app:frequentist}), meaning reduced prediction confidence when features are weakly correlated, as is desirable in general. An ideal method to weight features is one that can capture feature redundancy and balance it with feature relevance \cite{Peng-05}. This suggests a learning approach if possible, which, as we shall see, is unlikely to pick uniform or unit sum weights.

\subsection{Learning the weights}
\label{sec:learning}

Assuming a training dataset ${\cal D}=\{(x_k,\y_k), k=1,\ldots,N\}$, the most direct way to learn the composite weights is to maximize their likelihood under the composite predictive distribution (or, equivalently, minimize their ``cross-entropy''):
\begin{equation}
\label{eq:train_likelihood}
\max_{\blambda\succeq 0} U(\blambda),
\qquad \text{with} \quad
U(\blambda) \equiv\sum_{k=1}^N \log p_\blambda(x_k|\y_k).
\end{equation}

From~(\ref{eq:log_pool}), we see that the set of conditional distributions $p_\blambda(x|\y)$ spanned by the weights~$\blambda$ is the exponential family with natural parameter~$\blambda$ and basis functions given by the feature-based log-likelihoods:
$$
p_\blambda(x|\y) = \pi(x) \exp[\blambda^\top \bell(x,\y) - a(\blambda,y)],
$$
with:
$$
\ell_i(x,\y) = \log f(z_i|x),
\qquad
a(\blambda, \y) \equiv \log \sum_{x\in{\cal X}} \pi(x) e^{\blambda^\top \bell(x,\y)}.
$$

It follows that the utility function $U(\blambda)$ in~(\ref{eq:train_likelihood}) is concave as a general property of likelihood in exponential families, hence the learning can be implemented using a standard convex optimization scheme such as the limited-memory BFGS algorithm \cite{Byrd-95}. See Appendix~\ref{app:training} for some implementation details.

Because some positive weight constraints may turn out inactive, learning has a natural tendency to produce sparse weights. As illustrated in Figure~\ref{fig:disc_weight_plot} on the breast cancer UCI dataset\footnote{\url{https://archive.ics.uci.edu/ml/datasets/Breast+Cancer+Wisconsin+(Diagnostic)}}, there is no clear relation between optimal weights and feature-level discrimination power, showing that maximum likelihood learning differs from univariate feature selection as it implicitly takes feature redundancy into account via joint weight optimization.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=.6\textwidth]{disc_weight_plot.pdf}
  \end{center}
\caption{Distribution of optimal composite weights vs individual feature discrimination power in a binary classification task (breast cancer UCI dataset). Discrimination is measured by the maximum KL divergence between class-conditional generative distributions.}
\label{fig:disc_weight_plot}
\end{figure}

As is customary and generally good practice, the training examples may be weighted in~(\ref{eq:train_likelihood}) so that the empirical distribution of classes matches the prior~$\pi(x)$, enforcing balanced classes if the prior is uniform. Also, some regularization may be needed for stability: we shall in practice maximize $U(\lambda)-\alpha \|\blambda\|^2$, where $\alpha>0$ is a small damping factor.

Other training variants that we mention here but do not particularly recommend for interpretability arise from adding basis functions to the exponential family. For instance, class-dependent offsets may be added so as to learn the ``prior'' together with the composite weights. It is also possible to perform unconstrained optimization to achieve higher training likelihood, but this means that some composite weights are then negative, in violation of the monotonicity property mentionned in Section~\ref{sec:log_pool}.

\paragraph{$I$-projection interpretation.}

%D(p\|p_0) = \sum_{x\in{\cal X}}\sum_{\y\in{\cal Y}} p(x,\y) \log \frac{p(x,\y)}{p_0(x,\y)},
Exponential family properties also imply that learning via likelihood maximization~(\ref{eq:train_likelihood}) is dual to an $I$-projection \cite{Csiszar-84}: 
\begin{equation}
\label{eq:maxent}
\min_{p\in{\cal P}} D(p\|p_0),
\qquad \text{with} \quad
p_0(x,\y) \equiv \pi(x)h(\y),
\end{equation}
subject to the constraints $p(\y)=h(\y)$ and $\E_p[\bell] \succeq \E_h[\bell]$, where $h(x,\y)$ is the joint empirical distribution of targets and data examples, $h(\y)$ is the corresponding marginal data distribution, and ${\cal P}$ is the set of joint probability distributions on ${\cal X}\times {\cal Y}$ with ${\cal Y}\equiv\{\y_1,\ldots,\y_N\}$. Note that the constraints implicitly approximate the data marginal and the mean log-likelihood values by their empirical estimates, confirming the need for regularization in practice.

The marginal constraint $p(\y)=h(\y)$ implies that only the conditional distribution $p(x|\y)$ is optimized, hence the $I$-projection is equivalent to maximum conditional entropy \cite{BergerA-96}. The mean log-likelihood constraints $\E_p[\bell] \succeq \E_h[\bell]$ encapsulate dependences between the target and the data, and consider a model admissible if it guarantees the same average log-likelihood levels as observed separately for each feature. This is weaker a constraint than assuming that the feature generative models $f(z_i|x)$ are true: it only assumes that the {\em description length} \cite{Grunwald-07} achieved by each feature model is a truthworthy upper bound.

%In the $I$-projection framework, the composite weights are the Lagrange multipliers associated with the mean log-likelihood inequality constraints.

An insightful analysis given in \cite{Grunwald-04} shows that the $I$-projection is, under broad conditions, a minimax strategy for prediction according to the log loss. Hence, any $I$-projection is in some sense a best possible predictive model given some arbitrary knowledge exctracted from the training data, which is represented by mean-value constraints and is inexact due to the limited training set size. Optimized composite likelihood only differs from other maximum entropy classifiers such as multinomial logistic regression by the underlying knowledge, which in this case may be interpreted in terms of maximum feature description length.

\section{Super composite likelihood}
\label{sec:super}

So far, we have assumed that each agent is given a single weight in the opinion aggregation~(\ref{eq:log_pool}). Let us now consider a more general pooling operator where agents can be weighted depending on the target variable:
\begin{equation}
\label{eq:super_pool}
p_\lambda(x|\y) \propto \pi(x) \prod_{i=1}^n \left[\frac{\ell_i(x)}{\ell_i(x_0)}\right]^{\lambda_i(x)},    
\end{equation}
where $x_0$ is some fixed target value, which we call the reference. As discussed in Appendix~\ref{app:super}, the main reason for introducing the normalization by a reference is to make the pooling externally Bayesian if the weighting functions~$w_i(x)$ sum up to one uniformly,
$$
\forall x\in{\cal X},\quad
\sum_{i=1}^n w_i(x) = 1,
$$
hence providing a sensible default rule to tune the weights. Note that (\ref{eq:super_pool}) amounts to a log-linear pool if the weights are target-independent as the reference is then effectless. Using the same analogy with Bayes rule as in Section~\ref{sec:bayes_rule}, we see that the quantity:
\begin{equation}
\label{eq:super_comp_lik}
L^{sc}_\blambda(x) \equiv 
\prod_{i=1}^n \left[\frac{\ell_i(x)}{\ell_i(x_0)}\right]^{\lambda_i(x)}
\end{equation} 
plays the role of a standard likelihood at prediction time. We call this form super composite likelihood (SCL) owing to the ``doubly composite'' aspect of combining distinctive feature-target pairs. SCL further generalizes CL since $L^{sc}_\blambda(x)=L^c_\blambda(x)/L^c_\blambda(x_0)$ for target-independent weights $\lambda_i(x)\equiv \lambda_i$, hence it is in this case proportional to CL and yields the same predictive distribution.

Conversely, when each target assigns a weight one to a single feature and zero to all the others, SCL turns out to be identical to the {\em class-specific} likelihood surrogate derived in \cite{Baggenstoss-03} from a generative model selection argument -- namely, the ``PDF projection theorem'', which in fact boils down to an $I$-projection \cite{Minka-04}. While our derivation stems from a {\em discriminative} model selection approach, it is interesting to see that it includes the class-specific method as a special case, and thus sheds new light on this method.

%Case of a single nonzero agent per feature. Then SCL is equivalent to the class-specific likelihood derived in \cite{Baggenstoss-03} based on a generative model selection argument  Our derivation is much different but interesting to see that we end up with a more general form.

A compelling advantage of SCL over CL is that it can deal with missing likelihood values, which happen if the feature generative distributions $f(z_i|x)$ are unknown for some target values~$x$. It is possible to assign zero weights $\lambda_i(x)$ to every feature-target pair for which $\ell_i(x)$ is unknown, yet using nonzero weights for other pairs. Likewise, the features for which the reference likelihood $\ell_i(x_0)$ is unknown receive zero weight for all target values, implying that they have no influence on the predictive distribution.

%Interest? If some feature likelihood values are missing, i.e. agents are not able to model the whole set of targets. It's a bit like censoring. Then it's ok if they can model at least two targets including the reference: $\ell_i(x)$ is unknown for the other values but we can give them zero weight so it doesn't matter.

If training data is available, the SCL weights may be learned by maximum likelihood as in Section~\ref{sec:learning}, yielding a similar convex problem similar. Since SCL enables to explore a wider class of predictive models than CL, it has the potential to achieve better prediction performance as long as overfitting does not happen. The $I$-projection interpretation still holds, yet with class-specific mean-value constraints:
$$
\forall a\in{\cal X}, \forall i=1,\ldots,n,
\quad
\E_p\left[\delta_{xa}\log \frac{f(z_i|x)}{f(z_i|x_0)}\right]
\geq \E_h\left[\delta_{xa}\log \frac{f(z_i|x)}{f(z_i|x_0)}\right],
$$
where $\delta$ denotes the Kronecker delta.
%$$
%\forall x\in{\cal X}, \forall i=1,\ldots,n,
%\quad
%p(x) \E_{f(z_i|x)}\left[\log \frac{f(z_i|x)}{f(z_i|x_0)}\right]
%\geq h(x) \E_{h(z_i|x)}\left[\log \frac{f(z_i|x)}{f(z_i|x_0)}\right]
%$$

%Another shortcoming is that SCL is not guaranteed to be asymptotically consistent, see Appendix~\ref{app:frequentist}, so it may yield biased MCLE despite

%Disadvantage: interpetation is relative to an arbitrary class $x_0$. Advantages: richer model (so better predictions), probabilty ratios stable to class addition. Drawbacks: require a reference, no more consistency. Still interpretable, that's why it is worthwile mentioning.

%Argue that SCL more robust than CL to mispecification of feature models. Intuitively, if a feature model is poor, the empirical description length is large, hence can be achieved with a `small' Lagrange multiplier $\lambda_i(x)$. Therefore, unreliable likelihoods $\ell_i(x)$ tend to get low weights.

% which does not contradict the axiomatic derivation in \cite{Genest-86b} as it restricted itself to operators that only depend on the probas evaluated at~$x$ unlike~(\ref{eq:super_pool}).

%The real motivation: when agents operate on restricted target sets. Without loss of generality, assume binary sets. Trick: for each such agent, say we have $p(z_i|x_1)$ and $p(z_i|x_0)$. We define $p(z_i|x)=p(z_i|x_0)$ if $x\not= x_1$. It does not make any sense but the likelihood ratio is then one for any $x\not= x_1$ so won't hamper the system.

%Due to the distribution of weights between clues, a drawback of composite likelihood is that it is prone to {\em information overload} in the sense that it tends to ``flatten out'' when too many clues are included as relevant clues then get downweighted. If one decides not to merge clues for computational reasons (this would require handling their joint distribution), one could hope to mitigate information overload by assigning strong weights only to those clues that are believed to be ``most informative''. 
% When chosen for computational simplicity, features may not only convey limited information at individual level, their informativeness may also be very much class-dependent. Consider, for instance, diagnosing a disease from a routine medical checkup. Body temperature may point to a bacterial infection by comparison with normality, but would not help detecting a non-infectious cardiovasc ular disease -- and conversely for, say, blood pressure.

%Reference is implicit if weights are class-independent as it is then effectless. Make sense to have a reference as it gives a customizable baseline for detection. Can be the feature distribution averaged over all classes (not recommended I would say) or the distribution of a particular class if that makes sense (e.g. healthy population). In this case, the super composite likelihood is equivalent to the standard one in the two-class case.

%If some classes are merged or split, then the other classes keep the same weights (assuming $h(\y)$ is kept constant, so we need to make sure the prior is adapted accordingly)! So if a patient was found to have 10 times more probability to be MCI than healthy (and healthy is the reference class), and we later split AD population into, say, pure AD and mixed AD, the MCI probably ratio fact will remain.

%Super composite likelihood is essentially about combining binary versions of composite likelihood. It is analogous to the ``one-versus-rest'' strategy in multiple logistic regression.

%Still externally Bayesian if weights sum up to one in each class. Alternatively, we can train this model by maximum likelihood, yielding a similar optimization problem as in Section~\ref{sec:learning}.

%Maxent interpretation. It is an alternative to the strategy considered above, the mean log-likelihood constraints may be expressed for each class rather than being averaged with respect to the prior, leading to a more general predictive model that we call ``super composite likelihood'', in which the weights become class-dependent (see Appendix~\ref{sec:super}). This variant has the potential for higher prediction performance, but poses some conceptual difficulties, in particular that of having to define an arbitrary reference class.

%used to predict whether breast lumps are benign or malignant based on a set of features extracted from fine-needle aspirate images.

%The most discriminative features tend to get large weight, however some quite discriminative features may get low weight if strongly correlated with a more discriminative feature

%Non-linear means that features with average discrimination power may get low weight as they are superseded by the most discriminative features. But some features with low discrimination power may also get large weight, which does not mean that they are very influential in practice (see above). This is illustrated in Figure~\ref{fig:disc_weight_plot} with the UCI breast cancer dataset.

%{\color{red} Equality constraints (negative weights) are less interpretable.}

%{\color{red} Offsets.}

%{\color{red} Data weights.}

%Need to weight the data to adjust to prior as a general measure when minimizing cross-entropy, which should yield a conditional model that approximates the ``true'' conditional distribution, $q(x|y)\approx p(x|y)$. So the model will absorb the ``true'' marginal of $x$, which should thus be replaced by the prior if we want the model to mimic ideal Bayesian inference. In our case, it means that the composite likelihood approximates the true likelihood.

%We can further enforce exact prior matching by including offsets ($K-1$ free parameters), in which case the training phase learns $\pi(x)$ so that it matches the prior $\pi(x)$ in the sense $\int h(\y)p_\lambda(x|y)dy = \pi(x)$. This ``deforms'' the prior to improve the training score. But why is it important to enforce this constraint? Again, it all makes things just less interpretable.

%When including offsets, using homoscedastic generative variance estimation and equality constraints, composite likelihood is equivalent to binomial logistic regression. However, equality constraints are not easily interpretable.

%In some sense, composite weights replace the multiple testing correction required in the frequentist inference paradigm for proper control of false positives. Here, we don't control false positives, we control prediction accuracy.

%Why not using the feature-based posteriors or Bayes factors rather than the likelihoods? It's a possibility (perhaps a connection here with Gr\"unwald's luckiness). A justification for not doing it is that the coordinator should discard the priors used by the agents. The constraints are not purely evidence-based since the moments depend on the coordinator's prior but it is arguable a mess of priors wouldn't be good.

%Single feature case ($y=z$): the maxent solution has the form $p(x|y)\propto\pi(x)p(\y|x)^\lambda$. We expect to have $\lambda=1$ but it's not necessarily the case. It is if $h(\y)=\int\pi(x)p(\y|x)dx$ because then $h(\y)p(x|y)=\pi(x)p(\y|x)$ so the constraint is verified (and active). This also tells us that the data marginal should be consistent with the prior on labels for this much expected consistency property to hold. So, either weight datapoints so as to match a desired prior (e.g., uniform) or take the empirical distribution of labels as the ``prior''.

%Which brings another question: in this case, the (unique) weight is insensitive to the prior -- is it true in general? The answer is no. The weights generally depend on the prior. Hence the maxent composite likelihood is prior-dependent. This is an important conceptual difference with the classical notion of likelihood.

\section{Discussion}
\label{sec:discussion}

Beside greater flexibility than CL, a disadvantage of SCL it that interpretation becomes relative to an arbitrary reference~$x_0$. In some applications, there may exist a natural reference class to cling to in common practice (for instance, normality in medical diagnosis), but this is not always the case.

It is easy to build a good classifier. You just need to pick a parametric family that is big enough to produce a good fit yet small enough to avoid overfitting. Less easy is to interpret what the classifier does. Deep learning is cool but it is relevant to AI only if the representations it learns are generic enough to be used in other problems, i.e. it should ultimately be validated via transfer learning.

{\color{red} Why prediction performance is still important? Because it validates (or not) interpretation. Interpretation is the goal but prediction performance is a crucial intermediate step.}

{\color{red} We cannot rank features according to their weights. A feature might incidentally get a large weight because its likelihood is flat. Feature relevance also depends on conditional likelihood distribution. Consider mutually independent features: each one recieves a unit weight, but some may be independent from target.}

{\color{red} Can we use the same dataset to learn both the feature generative distributions and the composite weights? In principle, nothing speaks against this.}

{\color{red} Semi-generative method. Good for shallow learning or transfer learning. Still need representation learning methods.}

{\color{red} Alternative to logistic regression in deep learning? Would require a game-theoretic learning approach... future work.}

{\color{red} Composite Bayes is invariant by affine transformation of the data (unlike multinomial LR, which is problematic when damping is applied).}

{\color{red} Is damping important? (it guarantees training cost is stricly convex).}

{\color{red} Can we interpret maximum weight likelihood as a ``maximum relevance minimum redudancy'' feature selection approach?}

Is it new to see composite likelihood as opinion pooling? Yes it is. Some previous attempts were made to derive composite likelihood as some sort of consensus generative model \cite{Wang-14}. Also some previsous attemps at Bayesian inference from CL using asymptotic theory.

%In other words, is $\bar{a}(\blambda)$ a measure of feature redundancy? At fixed $\y$, if one $\lambda_i\to\infty$, then $a(\blambda,\y)\ot\infty$ except if $\ell_i(x,\y)\preceq 0$, in which case it goes to $-\infty$. If using likelihood ratios, this only happens if the reference class is the most likely for $\y$. But then we can prove that it can't happen for all $\y$ (otherwise the associated moment would be negative), so basically $\bar{a}(\blambda)$ blows up when at least one weight gets large. And if several weights get large simultaneously, then it blows up even more so we might consider it a redundancy measure to some extent...

Some experimental observations. In the iris dataset (historical data used by Fisher to demonstrate linear discriminant analysis), composite Bayes finds that petal length is the most relevant feature followed by petal width, while sepal length and width get zero weight. In the wine dataset, the flavanoids is the most relevant feature for standard composite Bayes, but it's color intensity closely followed by flavanoids for the super composite version. In the breast cancer, both approaches are equivalent since it is a binary problem, weights are very sparse: only 13/30 are non-zero with worst area the most relevant. In the digits prolem, the most relevant features tend to be in the image center; composite Bayes can achieve similar performance to multinomial logistic regression with about 10 times less parameters.

Number of independent discriminative parameters in logistic regression: $(K-1)(F+1)$. In standard composite inference: $F$ (or $K-1+F$ with offsets). Difference $=(K-2)(F+1)+1$ (or $=(K-2)F$). In super-composite inference: $(K-1)(F+1)$ (if reference is a class). Number of generative parameters in composite inference using univariate Gaussian models: $2K$ with heteroscedasticity, $K+1$ with homoscedasticity.

Intuition behind weights: notion of global feature relevance. But does $\lambda_1>\lambda_2$ imply that $z_1$ is more relevant than $z_2$ for a particular instance? No. But we can assess the influence of each scaled likelihood in order to rank features according to their influence on the decision.

CL is a concept from computational statistics that has mainly been developed so far in a frequentist perspective as a surrogate for the maximum likelihood method. We have shown a deep connection between CL and probabilistic inference, thereby establishing CL as a class of discriminative models. Because CL is built from a set of marginal feature-specific generative distributions, it is in essence a two-step semi-generative, semi-discriminative learning approach. In the first, ``generative'' phase, the feature distributions are learned; in the second, ``discriminative'' phase, the feature weights are learned. This strategy can be thought of as a form of non-adaptive boosting.

%The first phase corresponds to the training of ``weak learners''. The second phase amounts to a form of boosting.

A purely discriminative learning could be used instead but...
Why potentially more efficient than pure discriminative training: because generative training is always more efficient than discriminative training if models are comparable. So the key point is that ``we have less parameters in the discriminant phase". Good in small datasets. But also in asymptotic regime if the features are weakly or highly correlated (???).

Logistic regression / naive Bayes example. Under homoscedasticity (assuming that each feature has class-independent variance), CBI is equivalent to logistic regression -- because the predictive distribution family is the same. This is a case where BCI brings nothing. But consider heteroscedasticity, then BCI yields a quadratic classifier. Compared to a fully discriminative model, the number of parameters to learned in the discriminative phase is reduced by half.

The first training phase is easier if supervised but could be unsupervised too (using mixture models). How do we then deal with label switching issues? Can we safely assume conditional feature independence {\em in the generative training phase}? I believe so provided that the marginal distribution parameters are disjoint. It's obvious in the supervised learning scenario.

If unsupervised, the first learning phase could be compared with contrastive pre-training of RBMs \cite{Hinton-06,Fischer-14}, which also optimize parameters for generation of observable features. BCI is comparable with an RBM with a single output unit (which is just a generative model assuming conditionial feature independence). The key difference is that the RBM is a full generative model while BCI only deals with marginal models, hence relying on weaker assumptions. This won't change anything in the pre-training phase but RBMs have to deal with more parameters in the generative learning phase. In fully supervised context, RBM pre-training is pointless since all parameters learned in the first phase will be overwritten. In BCI, pre-training is crucial even in supervised context because the parameters learned in the discriminative phase (the feature weighs) describe a sub-manifold of the predictive distribution family.

Needs features. Not a representation learning method, but could be coupled why not.

%{\color{red} Moreover, while RBM parameters are typically refined in a supervised discriminative learning step, disjoint set of parameters for SCL. Dunno how to say that.}

%In such context, SCL competes with classical discriminative models (logistic regression, Gaussian processes \cite{Rasmussen-06}, maximum entropy models \cite{BergerA-96}, etc.), and may compare more or less favorably in practice depending on the amount of training data. For relatively small training datasets, we may hope for more accurate inference using~SCL than using traditional discriminative models, extrapolating from the results of \cite{Ng-01} regarding the comparison between logistic regression and na\"ive Bayes classifiers.
%{\color{red}Ici, il manque la comparaison avec les RBMs qui ne sont pas (forcement) des modeles discriminatifs.}

%%, hence alleviating the need for heavily supervised model training

%In summary, CL has the potential to yield weakly supervised or unsupervised Bayesian-like inference procedures depending on the particular task at hand. This property reflects the encoding of statistical relationships between the data and {\em all} unknown parameters. CL thus appears as a trade-off between generative models, which are optimal for unsupervised learning but possibly intractable, and traditional discriminative models (logistic regression, Gaussian processes \cite{Rasmussen-06}, maximum entropy models \cite{BergerA-96}, etc.), which are inherently supervised. CL models are discriminative models assembled from atomic generative models and, from this point of view, may be considered as {\em semi-generative} models.

%CL may be considered as a {\em semi-generative} model: a discriminative model assembled from partial generative models.

%As a note of caution, we shall stress that the pre-determined weights assigned to the different associations between observed and unobserved values represent prior knowledge regarding the informativeness of clues. A poor choice of weights will inevitably result in a poor approximation to the ``true'' Bayesian posterior -- the posterior that would be obtained from a realistic generative model if it was tractable. In future work, we will investigate feature selection strategies to mitigate this problem.

% Improve the discussion on following aspects:
% ** Why is it compatible with unsupervised learning? Give more insight.
% ** Stress the contribution: class-specific weighting.
% * Pivotality argument.
% * Bayes is a special case of composite Bayes.

{\color{red}Product of fucking experts \cite{Hinton-02}. Log-linear pool to build a generative model (assuming in fact independence between experts). Better seen as a ``log-mixture model''. Sounds like each expert is associated with a class, so it's the same as a RBM. Contrastive learning approximates ML parameter estimation. The experts are learning jointly. It's just a generative model.}

{\color{red}Two-step training: generative phase then discriminative phase. Why is it cool?}

{\color{red}CBI is just a way to reweight Naive Bayes. What's the big deal? Can we really expect massively superior performance? Are we just talking about realistic credibility sets?}

Future work: can we collapse generative and discriminative training into a single step? Maybe using the concept of {\em minimally informative likelihood} \cite{Yuan-99b}.

\appendix

\section{Basic frequentist properties of composite likelihood}
\label{app:frequentist}

Let a multivariate observable variable~$\y \sim p(\y|x_\star)$ distributed according to some target value~$x_\star$. For any hypothetical target value~$x$ and any extracted feature $z_i\sim f(z_i|x)$, the following double inequality holds:
$$
0 \leq
E\left[
\log \frac{f(z_i|x_\star)}{f(z_i|x)}
\right]
\leq
E\left[
\log \frac{p(\y|x_\star)}{p(\y|x)}
\right],
$$
where the expectation is taken with respect to the true distribution $p(\y|x_\star)$. This fact follows from two basic properties of KL~divergence: positivity and partition inequality.

Using a weighted sum of such inequalities, we can bracket the expected variations of the logarithm of any composite likelihood function~(\ref{eq:comp_lik}) with positive weights~$\blambda\succeq 0$:
\begin{equation}
\label{eq:variation_bound}
0 \leq
E\left[ \log \frac{L_c(x_\star, \blambda)}{L_c(x,\blambda)} \right]
\leq 
s_\blambda E\left[ \log \frac{L(x_\star)}{L(x)} \right]
,
\end{equation}
with $s_\blambda\equiv \|\blambda\|_1 =\sum_i \lambda_i$, and $L(x)$ the true likelihood function, {\em i.e.} $L(x)\equiv \log p(\y|x)$.

This implies two general asymptotic properties of composite likelihood:
\begin{itemize}
\item {\em Consistency.} The expected composite log-likelihood is maximized by~$x_\star$.
\item {\em Conservativeness for unit sum weights.} If the weights sum up to one or less ($s_\blambda\leq 1$), composite likelihood ratios of the true target~$x_\star$ vs.~other target values~$x$ tend to be underestimated, {\em i.e.} are smaller on the average than corresponding true likelihood ratios.
\end{itemize}

What about SCL???

\section{Conditional composite likelihood}
\label{app:conditional}

As a straightforward  extension of marginal CL, each feature-based likelihood may be conditioned by an additional ``independent'' feature $z^c_i(\y)$ considered as a predictor of the ``dependent'' feature, $z_i(\y)$, yielding the more general form:
\begin{equation}
\label{eq:cond_feat_lik}
\ell_i(x) = f(z_i|x,z^c_i).
\end{equation}

Conditioning may be useful if it is believed that $z^c_i$ alone is little or not informative about $x$, but can provide relevant information when considered jointly with $x$, as in the case of regression covariates, for instance. Equation~(\ref{eq:comp_lik}) then amounts to conditional CL \cite{Varin-11}, a more general form of CL also including Besag's historical {\em pseudo-likelihood} \cite{Besag-74} developed for image segmentation.

\section{Composite likelihood training}
\label{app:training}

Using the same notation as in Section~\ref{sec:learning}, the likelihood function in~(\ref{eq:train_likelihood}) can be expanded as follows:
$$
U(\blambda) 
= \sum_{k=1}^N \left[
\log \pi(x_k) + \blambda^\top \bell(x_k, \y_k) - a(\blambda,\y_k)
\right]
$$

Maximizing $U(\blambda)$ is therefore equivalent to maximizing: 
$$
\psi(\blambda) \equiv \blambda^\top \bar{\bell} - \bar{a}(\blambda), 
$$
with:
$$
\bar{\bell} \equiv \frac{1}{N} \sum_k \bell(x_k,\y_k),
\qquad
\bar{a}(\blambda) \equiv \frac{1}{N} \sum_k a(\blambda,\y_k).
$$

Note that $\psi(\blambda)$ is nothing but the dual function associated with the maxent problem~(\ref{eq:maxent}).

Derivatives...
$$
\nabla\psi(\blambda)
= \bar{\bell} - \frac{1}{N} \sum_k \nabla a(\blambda,\y_k) 
$$

$$
\nabla\nabla^\top\psi(\blambda)
= - \frac{1}{N} \sum_k \nabla \nabla^\top a(\blambda,\y_k)  
$$

We have:
$$
\nabla a(\blambda,\y) = \E_{\blambda}(\bell|\y),
\qquad
\nabla \nabla^\top a(\blambda,\y) = \text{Var}_{\blambda}(\bell|\y)
$$
proving that $a(\blambda,\y)$ is convex in $\blambda$ for any $\y$, and in turn proves the concavity of $\psi$.

To see this, let:
$$
z(\blambda,\y) 
\equiv e^{a(\lambda,\y)}  
= \sum_{x\in{\cal X}} \pi(x) e^{\blambda^\top \bell(x,\y)}
$$

We have:
$$
\nabla a = \frac{\nabla z}{z}
$$

$$
\nabla\nabla^\top a 
= \left(
\frac{\nabla \nabla^\top z}{z} 
- \frac{\nabla z \nabla z^\top}{z^2}
\right)
$$

$$
\nabla z 
=
\sum_{x\in{\cal X}} \pi(x) \bell(x,\y)  e^{\blambda^\top \bell(x,\y)}
$$

$$
\nabla \nabla^\top z 
=
\sum_{x\in{\cal X}} \pi(x) \bell(x,\y) \bell(x,\y)^\top e^{\blambda^\top \bell(x,\y)}
$$

\section{Super log-linear pool}
\label{app:super}

Opinions: $p_i(x)$.

Log-linear pool:
$$
p(x)\propto \prod_i p_i(x)^{\lambda_i}
$$

What about having $\lambda_i(x)$ depend on the target?

Can we do: $p(x)\propto \prod_i p_i(x)^{\lambda_i(x)}$? Is it externally Bayesian?

Let the opinions be multiplied by some function $f(x)$. The new opinions read: $q_i(x)=f(x)p_i(x)/z_i$ with $z_i=\int f p_i$. So the pool gives:
$$
q(x) \propto p(x) f(x)^{\sum_i \lambda_i(x)} \prod_i z_i^{-\lambda_i(x)}
$$

So, it's not EB because $z_i$ may differ from one expert to the other.

It feels more reasonable to do:
$$
p(x) \propto \prod_i \left[\frac{p_i(x)}{p_i(x_0)}\right]^{\lambda_i(x)}
$$

Is it EB? We have:
$$
\frac{q_i(x)}{q_i(x_0)} 
= \frac{f(x)p_i(x)}{f(x_0)p_i(x_0)}
$$
because the $z_i$'s cancel out. Therefore,
$$
q(x) \propto p(x) \left[\frac{f(x)}{f(x_0)}\right]^{\sum_i \lambda_i(x)}
$$

So it's EB if the unit sum weight constraints are satisfied.

\section{Notes Genest}

External Bayesian: pooling and updating commute.
A pooling operator is compliant with the `likelihood principle' if the pool at $x$ depends on the opinions only through the probabilities $p_i(x)$.

Characterizaton of externally Bayesian pooling operators (EBPOs).

Any EBPO LP-compliant that does not depend on $x$ is a log-pool with unit sum weights:
$$
p(x) \propto \prod_i p_i(x)^{w_i}
$$

Any EBPO LP-compliant is a generalized log-pool with unit sum weights:
$$
p(x) \propto \pi(x) \prod_i p_i(x)^{w_i}
$$

Weights can be negative if domain of $x$ is finite.

To characterize more general EBPOs (not LP-compliant), they introduce an equivalence class. Two sets of opinions are equivalent if any ratio of two experts is the same up to a constant factor. It means that $q\sim p$ if there exists $f(x)$ such that  
$q_i(x) \propto f(x) p_i(x)$, $\forall i$.

Define an arbitrary member $p_{a1},p_{a2},\ldots$ in each equivalence class.

Any EBPO is of the form:
$$
T(p_1,p_2,\ldots) \propto \pi_a(x) v_a(x) \frac{p_1(x)}{p_{a1}(x)},
\qquad
v_a \geq \max (p_{a1},p_{a2},\ldots)
$$

\section{Log-likelihood ratio}

$$
\log r = \log \frac{N(x;\mu_1,\sigma_1)}{N(x;\mu_0,\sigma_0)}
$$

$$
\log N(x;\mu,\sigma) = -\frac{1}{2}\left[
\log(2\pi\sigma^2) + \frac{(x-\mu)^2}{\sigma^2}
\right]
$$

$$
\log r = -\frac{1}{2}
\left[
\log \frac{\sigma_1^2}{\sigma_0^2}
+ \frac{(x-\mu_1)^2}{\sigma_1^2} - \frac{(x-\mu_0)^2}{\sigma_0^2}
\right]
$$

It follows:
\begin{eqnarray*}
\log r 
 & = & 
-\frac{1}{2}\left[
\log \frac{\sigma_1^2}{\sigma_0^2}
+ \frac{\sigma_0^2(x-\mu_1)^2 - \sigma_1^2(x-\mu_0)^2}{\sigma_0^2 \sigma_1^2}
\right] \\
 & = & 
\frac{1}{2}\left[
\log \rho^2
+ \frac{(\rho^2-1)x^2 - 2(\rho^2\mu_0-\mu_1)x+ (\rho^2\mu_0 ^2-\mu_1^2)}{\sigma_0^2}
\right]
\end{eqnarray*}
with $\rho=\sigma_1/\sigma_0$. In the case of same variance ($\rho=1$), this simplies to an affine function of $x$ with offset given by the center of $(\mu_0,\mu_1)$:
$$
\log r = \frac{\mu_1-\mu_0}{\sigma^2} \left(
x - \frac{\mu_0 + \mu_1}{2}
\right) 
$$

In any case,
$$
E[\log r | H_1] = 
\frac{1}{2}
\left[
- \log \frac{\sigma_1^2}{\sigma_0^2}
+ \frac{\sigma_1^2 + (\mu_1-\mu_0)^2}{\sigma_0^2}
- 1
\right]
$$

Let $\delta = (\mu_1-\mu_0)/\sigma_0$. We have:
$$
E[\log r | H_1] = 
\frac{1}{2}
\left[
\delta^2
- \log \rho^2
+ \rho^2  - 1
\right]
$$

\bibliographystyle{abbrv}
\bibliography{cvis,stat,alexis}

%%\input{draft.biblio}

\end{document}