### 2020-03-31 notes/global_optimizer/sgd_ep.tex

\documentclass{article}

\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,subfigure}

\title{Point matching notes} 
\author{Alexis Roche} 
%\date{}

\begin{document}

\maketitle

\section{Newton method and expectation-propagation}

EP may be seen as a generalization of Newton's method.

\section{Gaussian updates}

Let $c(x)=K \exp[-\frac{1}{2} (x-\mu)^\top \Sigma^{-1} (x-\mu)]$ in canonical form and $g(x)$ such that $-\log g(x)=a + g^\top (x-\mu) + \frac{1}{2} (x-\mu)^\top H (x-\mu)$. Let us give the canonical expression of $c(x)g(x)$:
\begin{eqnarray*}
\log cg (x)
& = & 
\log K -\frac{1}{2} (x-\mu)^\top \Sigma^{-1} (x-\mu)
- a - g^\top (x-\mu) - \frac{1}{2} (x-\mu)^\top H (x-\mu) \\
& = & 
\log K -a - g^\top (x-\mu) - \frac{1}{2} (x-\mu)^\top (\Sigma^{-1} + H) (x-\mu) \\
& = & 
\log K -a - g^\top (x-\mu) - \frac{1}{2} (x-\mu)^\top \Sigma_n^{-1} (x-\mu),
\end{eqnarray*}
with the new covariance matrix:
$$
\Sigma_n = (\Sigma^{-1} + H)^{-1}
.
$$

To find the new mean, just cancel the gradient of the above expression:
$$
\nabla \log cg 
= 
- g - \Sigma_n^{-1} (x-\mu),
$$
yielding:
$$
\mu_n = \mu - \Sigma_n g
.
$$

Finally, let us evaluate the multiplicative constant. By definition,
$$
\log K_n = \log cg(\mu_n),
$$
hence, denoting $\delta=\mu_n-\mu=-\Sigma_n g$:
$$
\log K_n = \log K -a - g^\top \delta - \frac{1}{2} \delta^\top \Sigma_n^{-1} \delta
$$

$$
\log K_n = \log K -a + \frac{1}{2} g^\top \Sigma_n g.
$$

\end{document}