### 2021-01-13 cv/cv.tex
Update cv

\documentclass[english,10pt,a4paper]{scrartcl}

\usepackage{times}
\usepackage{fullpage}
\usepackage{hyperref} 
\usepackage{url}
\usepackage{tabularx}
\usepackage{multicol}
\usepackage{eurosym}
\usepackage{color}

\title{Alexis Roche}

\author{}
\date{}

\makeatletter
\newcommand{\matphi}{\boldsymbol{\Phi}}
\newcommand{\req}[1]{Equation~(\ref{#1})}
\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}

\newcommand{\heading}[1]{\vspace*{.75cm}\noindent{\Large{#1}}\vspace*{-.25cm}\newline\noindent\rule{\textwidth}{0.8pt}\vspace*{.25cm}}

\newcommand{\bighead}[3]{\noindent\begin{minipage}{.85\textwidth}{\textbf{#1}}\newline {\textbf{#2}}\end{minipage}\begin{minipage}{.15\textwidth}\begin{flushright}{\textbf{#3}}\end{flushright}\end{minipage}\newline\noindent\rule{.85\textwidth}{0.4pt}}

\newcommand{\head}[2]{{\noindent {\textbf{#1}} $\mid$ {\textbf{#2}}}}

\makeatother

\begin{document}

%\maketitle

\begin{flushright} 
{\huge {\textbf{Alexis Roche}}}\\
\begin{small}
+351 910 327 756 $\mid$ \url{alexis.roche@gmail.com}\\
\url{https://www.linkedin.com/in/alexis-roche-3815592}\\
%French National $\mid$ DOB 12/8/1972 $\mid$ Married, 2 Children\\
French National $\mid$ Married, 2 Children\\
\end{small}
\end{flushright}

\heading{Professional Profile}

\noindent Scientific team leader with a specialist expertise in computer vision, data science and machine learning, who has operated in the fields of academic research, healthcare, electro-domestic and entertainment industry. Graduated from \'Ecole Centrale Paris, and holder of a PhD in Engineering Science. Strong experience in the development of both open source and commercial software. {\em Keywords:} artificial intelligence, image processing, computer vision, medical image analysis, machine learning, deep learning, statistics, Python,~C.

\heading{Key Skills}

\noindent \textbf{DATA SCIENCE:} In-depth knowledge of statistics, computer vision, medical image analysis, traditional machine learning and deep learning.

\ \\
\noindent \textbf{SCIENTIFIC COMPUTING:} Over ten years experience in Python programming using the NumPy/SciPy ecosystem and interfacing C~language with Python.

\ \\
\noindent \textbf{SCIENTIFIC WRITING:} Main author of 25 peer-reviewed scientific articles published in international journals, proceedings and books, co-author of more than 100 scholarly publications and 5 published international patents.

\ \\
\noindent \textbf{TECHNICAL LEADERSHIP:} Proven ability to coordinate a team and manage objectives in multidisciplinary environments.

%\ \\
%\noindent \textbf{NEUROIMAGING:} Extensive experience in developing algorithms for structural and functional brain magnetic resonance imaging (MRI) in the form of open-source and commercial software for neuroscience, clinical research \& clinical practice.

\heading{Career Summary}

\bighead{HEAD OF COMPUTER VISION \& ARTIFICIAL INTELLIGENCE}{Didimo, Porto, Portugal}{2019--present}
\nopagebreak
\begin{itemize}
\item Developing computer vision and machine learning algorithms for reconstruction of high-fidelity digital humans from photos.
\end{itemize}
\ \\

\bighead{SENIOR COMPUTER VISION SCIENTIST}{CoVii, Ar\c celik/Beko group, Porto, Portugal}{2017--2019}
\nopagebreak
\begin{itemize}
\item Developed embedded object recognition and tracking algorithms for smart domestic appliances using traditional computer vision and deep learning (VUXHub and Artisan applications demonstrated at IFA Berlin, 2017--18).
\end{itemize}

\ \\
\bighead{LEAD CLINICAL RESEARCH -- ADVANCED CLINICAL IMAGING TECHNOLOGY}{Siemens Healthineers / Lausanne University Hospital (CHUV), Switzerland}{2011--2017}
\nopagebreak
\begin{itemize}
\item Led algorithmic development of the {\em MorphoBox} Siemens prototype for brain morphometry using anatomical MRI to help radiological reading for patients with suspected neurodegeneration.
\item Contributed a number of clinical validation studies of brain morphometry in collaboration with neurologists and radiologists, via statistical analysis and data mining techniques, in Alzheimer's disease and other dementia, mild cognitive impairment, multiple sclerosis, migraine, anorexia, \ldots 
\item Co-supervised three PhD students and two MSc students.
\item Main organizer of annual Siemens/CHUV {\em brain imaging} workshops in Lausanne, Switzerland, 2012--2016.
\end{itemize}

\ \\
\bighead{PERMANENT RESEARCHER -- NEUROIMAGING INSTITUTE}{French Atomic Commission (CEA), Paris, France}{2002--2011}
\nopagebreak
\begin{itemize}
\item Developed advanced algorithms for image processing and statistical analysis of functional, anatomical and diffusion-weighted MRI data (population analysis, spatio-temporal image registration, real-time imaging).
\item French ANR-funded project management: PI (Karametria project, statistical analysis of brain structures, budget: 620 K\euro, 2009--2011), team leader (NIBB project, studying language in infants via functional neuroimaging, 2006--2009).
\item Active contributor to the NiPy software library (Neuroimaging in Python, \url{www.nipy.org}) from 2006.
\item Main supervisor of two PhD students and four MSc students.
\item Academic Guest at Computer Vision Laboratory, Swiss Federal Institute of Technology Zurich (ETHZ), Switzerland, 2009--2011.
\end{itemize}

\ \\
\bighead{POST-DOCTORAL RESEARCHER -- WOLFSON MEDICAL VISION LABORATORY}{University of Oxford, UK}{2001--2002}
\nopagebreak
\begin{itemize}
\item Developed advanced medical image registration algorithms.
\item Consulting for Mirada Solutions Ltd (now Siemens Molecular Imaging) on image-based deformation tracking (brain, liver).
\end{itemize}

\ \\
\bighead{CONTINGENT SCIENTIST -- FRENCH NATIONAL SERVICE}{General Directorate for Armament (DGA), Vernon, France}{1996--1997}
\nopagebreak
\begin{itemize}
\item Developed plane trajectory simulator using stochastic process models.
\end{itemize}

\heading{Education}
\nopagebreak
\head{PhD, Engineering Science (1997--2001)}{French National Research Institute (INRIA) \& University of Nice-Sophia Antipolis, France}
\begin{itemize}
\item Development of multimodal image registration algorithms, with applications in radiotherapy, image-guided surgery and neuroscience. With highest honor.
\end{itemize}

\ \\
\head{MSc, Cognitive Science (1995--1996)}{University Pierre \& Marie Curie, Paris \rom{6}, France}
\nopagebreak
\begin{itemize}
\item Internship at Experimental Psychology Laboratory, National Center for Scientific Research (CNRS), Paris, France, on modeling human perception of tempo using artificial neural networks.
\end{itemize}

\ \\
\head{Engineer Degree (equivalent MSc) (1993--1996)}{Ecole Centrale Paris (now CentraleSupelec), France}
\nopagebreak
\begin{itemize}
\item Third year specialization in Applied Mathematics.
\end{itemize}

%%\item Other training: Neuronal dynamics open course, EPFLx, Switzerland (2014).

\heading{Additional information}
\nopagebreak
\noindent 
\begin{minipage}{.15\textwidth}
\textbf{Languages:}
\end{minipage}
\begin{minipage}{.85\textwidth}
Native French, fluent English, basic Italian \& Portuguese. %German.
\end{minipage}

\ \\
\noindent 
\begin{minipage}{.15\textwidth}
\textbf{Programming:}
\ \\ 
\ \\
\ \\ 
\ \\
\end{minipage}
\begin{minipage}{.85\textwidth}
{\tiny$\bullet$} Languages: C, Python, Matlab, R. \\
{\tiny$\bullet$} Python packages: numpy, scipy, pylab, skimage, opencv, sklearn, pytorch, TensorFlow, keras.\\
{\tiny$\bullet$} C/Python integration via cython.\\
{\tiny$\bullet$} Version control software: git, svn, perforce. \\
\end{minipage}

\ \\
\noindent 
\begin{minipage}{.15\textwidth}
\textbf{Publications:}
\ \\ 
\ \\
\ \\
\ \\
\end{minipage}
\begin{minipage}{.85\textwidth}
{\tiny$\bullet$} 115 scholarly publications.\\
{\tiny$\bullet$} Full list available at ORCID: \url{http://orcid.org/0000-0002-4821-6893} \\
{\tiny$\bullet$} Google scholar statistics: 6225 citations, h-index: 32, i10-index: 63 (December 2020).\\
{\tiny$\bullet$} 5 published international patents. \\
\end{minipage}

%\ \\
%\noindent 
%\begin{minipage}{.15\textwidth}
%\textbf{ Reviewing:}
%\ \\ 
%\ \\
%\end{minipage}
%\begin{minipage}{.85\textwidth}
%IEEE Transactions on Medical Imaging, Medical Image Analysis, NeuroImage, Frontiers in
% Neuroscience, IEEE Transactions on Image Processing, Signal Image and Video Processing, MICCAI conference, \ldots
%\end{minipage}

\ \\
\noindent 
\begin{minipage}{.15\textwidth}
\textbf{Teaching:}
\ \\ 
\ \\
\  \\
\end{minipage}
\begin{minipage}{.85\textwidth}
{\tiny$\bullet$} Brain morphometry post-graduate lecture (2h), Lausanne: UNIL (2014/16), EPFL (2015/16).\\
{\tiny$\bullet$} Statistical analysis in neuroimaging, INSERM/CNRS continuing education, France (10h/year from 2003 to 2015).\\
\end{minipage}

\ \\
\noindent 
\begin{minipage}{.15\textwidth}
\textbf{References:}
\end{minipage}
\begin{minipage}{.85\textwidth}
Available on request.
\end{minipage}

%\input{biblio.tex}

\end{document}