### 2023-01-19 papers/robayes/notes.tex
Add notes about robust Bayes

\documentclass{article}

\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,subfigure}
%\usepackage[utf8]{inputenc}

\def\x{\mathbf{x}}
\def\y{\mathbf{y}}
\def\e{\mathbf{e}}
\def\th{{\boldsymbol{\theta}}}
\def\n{{\boldsymbol{\nu}}}
\def\u{\mathbf{u}}
\def\m{\mathbf{m}}
\def\v{\mathbf{v}}
\def\f{\mathbf{f}}
\def\E{\mathbb{E}}
\def\g{\mathbf{g}}
\def\h{\mathbf{h}}
\def\A{\mathbf{A}}
\def\B{\mathbf{B}}

\title{Robust Bayes Reloaded}
\author{Alexis Roche}
%\date{}

\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\maketitle

\section{Initial toughts}

The Bayesian paradigm is semi-generative in essence in that it relies on a parametric distribution model that incorporates a priori on unknown parameters. This a priori is non-testable in essence, and can thus be chosen in an arbitrary manner, which may lead to really poor inference models for high-dimensional models... This is an undeniable {\em weakness} of classical Bayesian analysis.

Robust Bayes analysis can be used to work around this problem. We may see it as a minimax risk strategy (instead of Bayesian) to select a reasonably low-dimensional parametric model to fight the curse of dimensionality. It is, in essence, a dimension reduction scheme.

Note that we are not talking about overfitting, at least in a traditional sense.

\bibliographystyle{plain}
\bibliography{alexis,stat}
\end{document}