### 2024-02-06 cv/cv.tex
Work on CV

\documentclass[english,10pt,a4paper]{scrartcl}

\usepackage{times}
\usepackage{fullpage}
\usepackage{hyperref} 
\usepackage{url}
\usepackage{tabularx}
\usepackage{multicol}
\usepackage{eurosym}
\usepackage{color}

\newcommand{\tabitem}{~~\llap{\textbullet}~~}

\title{Alexis Roche}

\author{}
\date{}

\makeatletter
\newcommand{\matphi}{\boldsymbol{\Phi}}
\newcommand{\req}[1]{Equation~(\ref{#1})}
\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}

\newcommand{\heading}[1]{\vspace*{.75cm}\noindent{\Large{#1}}\vspace*{-.25cm}\newline\noindent\rule{\textwidth}{0.8pt}\vspace*{.25cm}}

\newcommand{\bighead}[3]{\noindent\begin{minipage}{.85\textwidth}{\textbf{#1}}\newline {\textbf{#2}}\end{minipage}\begin{minipage}{.15\textwidth}\begin{flushright}{\textbf{#3}}\end{flushright}\end{minipage}\newline\noindent\rule{.85\textwidth}{0.4pt}}

\newcommand{\head}[2]{{\noindent {\textbf{#1}} $\mid$ {\textbf{#2}}}}

\makeatother

\begin{document}

%\maketitle

\begin{flushright} 
{\huge {\textbf{Alexis Roche}}}\\
\begin{small}
+351 910 327 756 $\mid$ \url{alexis.roche@gmail.com}\\
\url{https://www.linkedin.com/in/alexis-roche-3815592}\\
%French National $\mid$ DOB 12/8/1972 $\mid$ Married, 2 Children\\
French National $\mid$ Married, 2 Children\\
Based in Porto, Portugal.
\end{small}
\end{flushright}

\heading{Professional Profile}

\noindent Hands-on technical team leader, applied researcher and engineer at heart, PhD, with a specialist expertise in computer vision and artificial intelligence. Has operated in public research and private R\&D in various fields (healthcare, electro-domestic, entertainment). Strong experience in the development of both open source and commercial software. {\em Keywords:} applied mathematics, statistics, machine learning, computer vision, computer graphics, computer-aided diagnosis, smart appliances, digital avatars, Python, C.

\heading{Key Skills}

\noindent

\noindent \textbf{TECHNICAL LEADERSHIP:} Proven ability to coordinate a team and manage objectives in multidisciplinary environments.

\ \\
\textbf{COMPUTATIONAL STATISTICS:} 25 years experience developing inferential statistical models for machine learning, computer vision, computer graphics, \ldots

\ \\
\noindent \textbf{SCIENTIFIC PROGRAMMING:} 15 years experience programming in Python and~C language in the SciPy ecosystem.

\ \\
\noindent \textbf{SCIENTIFIC REPORTING:} Author of more than 100 peer-reviewed scientific articles published in international journals, proceedings and books, and 6 published international patents.

%\ \\
%\noindent \textbf{NEUROIMAGING:} Extensive experience in developing algorithms for structural and functional brain magnetic resonance imaging (MRI) in the form of open-source and commercial software for neuroscience, clinical research \& clinical practice.

\heading{Career Summary}

\bighead{HEAD OF ARTIFICIAL INTELLIGENCE}{Didimo, Porto, Portugal}{2019--present}
\nopagebreak
\begin{itemize}
\item Leading development of generation pipeline of high-fidelity avatars from photos. {\em Keywords:} facial reconstruction, deep face, animation retargeting, image processing. Accepted talk at SIGGRAPH'22.
\end{itemize}
\ \\

\bighead{SENIOR COMPUTER VISION SCIENTIST}{CoVii, Ar\c celik/Beko group, Porto, Portugal}{2017--2019}
\nopagebreak
\begin{itemize}
\item Developed embedded algorithms for smart domestic appliances (VUXHub and Artisan intelligent oven prototypes demonstrated at IFA Berlin, 2017--18). {\em Keywords:} image classification, object recognition \& tracking, deep learning.
\end{itemize}
\ \\

\bighead{LEAD CLINICAL RESEARCH -- ADVANCED CLINICAL IMAGING TECHNOLOGY}{Siemens Healthineers / Lausanne University Hospital (CHUV), Switzerland}{2011--2017}
\nopagebreak
\begin{itemize}
\item Led algorithmic development of computer-aided diagnosis software (released by Siemens Healthineers as part of the {\em AI-Rad Companion} solution) using brain morphometry for radiological reading. {\em Keywords:} magnetic resonance imaging (MRI), brain image segmentation, disease classification, neurodegeneration, Alzheimer's disease.
\item Main organizer of annual Siemens/CHUV {\em brain imaging} workshops in Lausanne, Switzerland, 2012--2016.
\end{itemize}

\ \\
\bighead{PERMANENT RESEARCHER -- NEUROSPIN (NEUROIMAGING INSTITUTE)}{French Atomic Commission (CEA), Paris, France}{2002--2011}
\nopagebreak
\begin{itemize}
\item Developed advanced algorithms for image processing and statistical analysis of brain imaging data. {\em Keywords:} functional, anatomical and diffusion-weighted MRI, image-based population analysis, spatio-temporal image registration, real-time imaging.
\item Funded project management: principal investigator ({\em Karametria} -- statistical analysis of brain structures, French National Research Agency project, budget: 620 K\euro, 2009--2011); team leader ({\em NIBB} -- studying language in infants via functional neuroimaging, French National Research Agency project, 2006--2009).
\item Active contributor to the NiPy software library (Neuroimaging in Python, \url{www.nipy.org}) from 2006.
\item Academic Guest at Computer Vision Laboratory, Swiss Federal Institute of Technology Zurich (ETHZ), Switzerland, 2009--2011.
\end{itemize}

\ \\
\bighead{POST-DOCTORAL RESEARCHER -- WOLFSON MEDICAL VISION LABORATORY}{University of Oxford, UK}{2001--2002}
\nopagebreak
\begin{itemize}
\item Developed advanced medical image registration algorithms.
\item Consulting for Mirada Solutions Ltd (now Siemens Molecular Imaging) on image-based deformation tracking (brain, liver).
\end{itemize}

\ \\
\bighead{PhD CANDIDATE -- EPIDAURE PROJECT}{French National Research Institute (INRIA), Sophia Antipolis, France}{1997--2001}
\nopagebreak
\begin{itemize}
\item Development of multimodal image registration algorithms, with applications in radiotherapy, image-guided surgery and neuroscience. 
\end{itemize}

\ \\
\bighead{CONTINGENT SCIENTIST -- FRENCH NATIONAL SERVICE}{General Directorate for Armament (DGA), Vernon, France}{1996--1997}
\nopagebreak
\begin{itemize}
\item Developed plane trajectory simulator using stochastic process models.
\end{itemize}

\heading{Education}
\nopagebreak
\head{PhD, Engineering Science (1997--2001)}{University of Nice-Sophia Antipolis, France}
\begin{itemize}
\item With highest honor.
\end{itemize}

\ \\
\head{MSc, Cognitive Science (1995--1996)}{University Pierre \& Marie Curie, Paris \rom{6}, France}
\nopagebreak
\begin{itemize}
\item Internship at Experimental Psychology Laboratory, National Center for Scientific Research (CNRS), Paris, France, on modeling human perception of tempo using artificial neural networks.
\end{itemize}

\ \\
\head{Engineer Degree (equivalent MSc) (1993--1996)}{Ecole Centrale Paris, France}
\nopagebreak
\begin{itemize}
\item Third year specialization in Applied Mathematics.
\end{itemize}

%%\item Other training: Neuronal dynamics open course, EPFLx, Switzerland (2014).

\heading{Additional information}
\nopagebreak
\noindent

\begin{tabular}{p{0.2\linewidth}p{0.8\linewidth}}
  \textbf{Languages:} & Native French, fluent English, basic Portuguese \& Italian.\\
  %\hline
  \ \\
  \textbf{IT skills:} &
  \tabitem Programming languages: Python, C, C++, Matlab, R.\\
  & \tabitem Long-term experience with scientific Python packages (numpy, scipy, pylab).\\
  & \tabitem CV/AI Python packages: skimage, PIL, opencv, sklearn, pytorch, tensorflow, keras.\\
  & \tabitem C/Python integration via Cython.\\
  & \tabitem Version control software: git, svn, perforce.\\
  & \tabitem Agile workflow (mainly Scrum) using JIRA \& Confluence.\\
  & \tabitem Reporting: \LaTeX, MS~Office, OpenOffice, LibreOffice, Google~Docs.\\
  %\hline
  \ \\
  \textbf{Publications:} &
  \tabitem ORCID: \url{http://orcid.org/0000-0002-4821-6893}.\\
  & \tabitem 116 scholarly publications.\\
  & \tabitem Google scholar statistics: 7603 citations, h-index: 35, i10-index: 74 (May 2023).\\
  & \tabitem 6 published international patents. 2 ongoing applications.\\
  %\hline
  \ \\
  \textbf{Academic work:} & 
  \tabitem Post-graduate teaching in computer vision and medical image analysis: about 100~hours (France: Ecole Centrale Paris, University of Nice, INSERM/CNRS continuing education; Switzerland: EPFL Lausanne, University of Lausanne).\\
  & \tabitem Long-time scientific journal reviewing experience (IEEE Trans.~Medical Imaging, Medical Image Analysis, NeuroImage, IEEE Trans.~PAMI, IEEE Trans.~Signal Processing, Frontiers in Neuroscience, \ldots).\\
  & \tabitem Supervision of two PhD students and four MSc students as main supervisor, three PhD students and two MSc students as co-supervisor.\\
  %\hline
  \ \\
  \textbf{References:} & Available on request.
\end{tabular}

%\input{biblio.tex}

\end{document}